---
title: "Pair Assignment 1"
author: "Tori Dykes and Mansi Wadhwa"
date: "Tuesday, October 04, 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(stargazer)  # for regression tables
library(ggplot2) # for plotting tables

```


The two datasets are: Swiss Fertility and Socio-economic indicators (1888) Data from R's built-in datasets and worldwide alcohol consumption (by country) from the fivethirtyeight datasets. 


## Swiss Fertility and Socio-economic indicators (1888) Data

```{r}
data(swiss)
names(swiss)
```

The first dataset 'swiss' is the Swiss Fertility and Socio-economic indicators (1888) Data from R itself that provides standardized fertility measure and socioeconomic indicators for each of the 47 French-speaking provinces of Switzerland. It contains a total of 47 observations and 6 variables (all in percent).

### Exploring the relationship between the variables Fertility and Catholic:

The variable Fertility gives the standardized fertility measure and the variable Catholic gives us the percentage of Catholics in the population of each province. Quite some research has been focussed on exploring the impact of socio-economic, cultural and religious conditions on fertility levels. The case of Switzerland is interesting due to its cultural diversity and the fact that its population declined greatly in 1885.
 
```{r}
summary(swiss$Fertility)
summary(swiss$Catholic)

hist(swiss$Fertility, main = 'Standardized measure of fertility for 47 Swiss provinces in 1888')
```

To understand the relationship between fertility and Catholic better, we create a factor variable 'factor' that creates 3 categories for the variable Catholic.Variable factor takes 3 levels 0 to 30, 30 to 70 to 100. Level 0 to 30 represents those provinces where the Catholic population is less than or equal to 30% and so on. 

```{r}
swiss$factor <- cut(swiss$Catholic, c(0, 30, 70, 100))
boxplot(swiss$Fertility ~ swiss$factor, main = 'Variation in fertility across provinces
\nwith different levels of Catholic population \n', xlab = 'Levels of Catholic population', 
ylab = 'Fertility indicator')

```

The boxplot shows that provinces with a Catholic population between 30 and 70 percent have a substantially lower ferility level than the other two categories. 

```{r}
cor.test(swiss$Fertility, swiss$Catholic)

```
The Pearson's correlation for this pair of variables is 0.46 and hence, shows a positive but moderately strong relationship.

We can also further explore the data by way of a simple regression exploring the relationship between fertility and the percentage of Catholics living in a province.

```{r, results='asis'}

fertilreg <- lm(swiss$Fertility ~ swiss$Catholic)

stargazer(list(fertilreg), header = F, float = F, single.row = T)

```

Finally, we can plot the values and apply a regression line (as well as shaded confidence intervals) to get a further visual perspective:


```{r}

ggplot(swiss, aes(x=Catholic, y=Fertility)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth(method=lm)

```

We see that while there is a positive relationship between the two variables, it's not very strong. Moreover, we see a strong bimodal distribution in the datapoints, with there being significant clustering around both very low rates and very high rates of Catholicism and few data points between them. 

## Worldwide Alcohol Consumption (by country)

This dataset, published by the WHO in 2010, looks at the amount of beer, spirits, or wine were consumed per person in a given country in 2010. The units (according to this post: http://fivethirtyeight.com/datalab/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) are individual servings -- so, a can of beer, a glass of wine, or a shot of a spirit. The dataset also looks at the total amount of alcohol consumed in liters. 

```{r}
alcoholconsump <- read.csv('drinks.csv', skip = 0, header = T, sep = ',',
              fileEncoding = 'latin1', encoding = 'UTF-8',
              stringsAsFactors = F)
```


Here are some general overview statistics for the data set:

&nbsp;

**Beer Consumption**

Average consumption per person per year (across all countries): `r mean(alcoholconsump$beer_servings)`

Max consumption per person per year: `r max(alcoholconsump$beer_servings)` by `r alcoholconsump$country[c(which.max(alcoholconsump$beer_servings))]`

**Wine Consumption**

Average consumption per person per year (across all countries): `r mean(alcoholconsump$wine_servings)`

Max consumption per person per year: `r max(alcoholconsump$wine_servings)` by `r alcoholconsump$country[c(which.max(alcoholconsump$wine_servings))]`

**Spirits Consumption**

Average consumption per person per year (across all countries): `r mean(alcoholconsump$spirit_servings)`

Max consumption per person per year: `r max(alcoholconsump$spirit_servings)` by `r alcoholconsump$country[c(which.max(alcoholconsump$spirit_servings))]`

**Overall Alcohol Consumption (pure)**

Average consumption per person per year (across all countries): `r mean(alcoholconsump$total_litres_of_pure_alcohol)`

Max consumption per person per year: `r max(alcoholconsump$total_litres_of_pure_alcohol)` by `r alcoholconsump$country[c(which.max(alcoholconsump$total_litres_of_pure_alcohol))]`


&nbsp;

Since there are 193 countries in this list, we can't look at everyone at once. To give a bit more perspective, here are some visualizations just for the top 10 countries in each category:

```{r, echo=FALSE}

BEERalcoholconsump <- alcoholconsump[order(-alcoholconsump$beer_servings),]
top10beer <- BEERalcoholconsump[c(1:10),]

barplot(top10beer$beer_servings, names.arg = top10beer$country,xlab = "", ylab = "Beer Servings",col = "gray",
main = "Top 10 Consumers of Beer",border = "black", las=2)

```

```{r, echo=FALSE}

# need to troubleshoot wonkiness of Czech Republic in graph above

# ggplot(top10beer, aes(x = top10beer$country[order(top10beer$beer_servings)], y = top10beer$beer_servings)) + geom_bar(stat = "identity") --> keeping this here for future reference; Tori can't get this to work properly but maybe will troubleshoot it later

```

```{r, echo=FALSE}

WINEalcoholconsump <- alcoholconsump[order(-alcoholconsump$wine_servings),]
top10wine <- WINEalcoholconsump[c(1:10),]

barplot(top10wine$wine_servings, names.arg = top10wine$country,xlab = "", ylab = "Wine Servings",col = "gray",
main = "Top 10 Consumers of Wine",border = "black", las=2)

```


```{r, echo=FALSE}

SPIRITalcoholconsump <- alcoholconsump[order(-alcoholconsump$spirit_servings),]
top10spirit <- SPIRITalcoholconsump[c(1:10),]

barplot(top10spirit$spirit_servings, names.arg = top10spirit$country,xlab = "", ylab = "Spirit Servings",col = "gray",
main = "Top 10 Consumers of Spirits",border = "black", las=2)

```

```{r, echo=FALSE}

PUREalcoholconsump <- alcoholconsump[order(-alcoholconsump$total_litres_of_pure_alcohol),]
top10pure <- PUREalcoholconsump[c(1:10),]

barplot(top10pure$total_litres_of_pure_alcohol, names.arg = top10pure$country,xlab = "", ylab = "Consumption (Liters)",col = "gray",
main = "Top 10 Consumers of Pure Alcohol",border = "black", las=2)

```